---
title: "Assignment 3 - Causal inference"
author: "Sofie Ditmer"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Semester4_Assignment3")
library(pacman)
p_load(tidyverse, ggplot2, brms, dplyr)
```

## Assignment 3 - Exploring causal inference issues

In this assignment we explore some issues related to multiple regressions (regressions with more than one predictor), and inferred (causal) relations between variables. N.B. the data is simulated (to make sure I know the actual mechanism generating it), but it's based on a real study. So bear with a longish introduction to get into the details of what we are doing and why it is important.

### Altercentric intrusion in schizophrenia

People with schizophrenia often report altered control and distinction of self-other representations: intrusive thoughts, hearing of voices, delusions of mind reading, paranoia, etc (a substantial portion of the psychotic symptoms experienced in schizophrenia). These have been variously attributed to hypermentalizing (over attribution of mental states to others), social impairment (over preoccupation with own thought processes), hyper socialization (inability to inhibit information from others), etc.

The current study investigates 
1) whether schizophrenia is indeed related to altered control and distinction of self-other representations, in particular altercentric intrusions (inability to inhibit social information)
2) whether these are related to the relevant psychotic symptoms. N.B. the actual study also investigates egocentric intrusion, do check the papers below if interested.

The task is a slightly modified version of this: https://www.ncbi.nlm.nih.gov/pubmed/20731512 You look at a picture with some dots visible to you, as well as with a different person with a different set of dots visible to them. The number of dots you see and that the other sees can be the same (congruent condition) or not (incongruent condition). You are tasked to indicate whether a given number (e.g. 3) matches the number of dots you see (and the dots visible to the other person are irrelevant to the task).


The tasks investigates altercentric intrusion: will your reaction time change according to whether the other person is seeing the same amount of dots as you, or not? The idea is that if you correctly inhibit social information, your reaction time should not change, as the information about the other person is not relevant. On the contrary, if you nevertheless use task irrelevant social information, you'll be slower at indicating whether 3 is the right number of dots when the other person sees a different amount of dots than you (conflicting information).
The bigger the difference between RTs in the congruent and incongruent condition the bigger the altercentric intrusion effect.

For each participant you have 6 variables: 
1) ID
2) AltercentricIntrusion (continuous score)
3) Diagnosis (schizophrenia vs. control)
4) VoiceHearing (severity of voice hearing symptoms, continuous score of the severity of the symptom as measured by a clinician)
5) MindReading (severity of delusions of mind reading, continuous score of the severity of the symptom as measured by a clinician)
6) Apathy (severity of lack of motivation in taking care of oneself, from washing to showing up at work, continuous score of the severity of the symptom as measured by a clinician).

The research questions you have to answer are the following:

## First part

Q1.1) Does schizophrenia involved altercentric intrusion? Define model and priors. Test the implications of your priors (prior predictive checks) and if needed adjust them. Run the model. Test the quality of the fitted model (posterior predictive checks). Assess the evidence in favor of an increased altercentric intrusion in schizophrenia. Report the model and the results, including plots.

```{r}
# Load the data
d <- read_csv("data.csv")
summary(d)

# We relevel the factors - instead of 0 and 1 we want them to be called "Controls" and "Schizophrenia"
d$Diagnosis <- plyr::revalue(as.character(d$Diagnosis), 
                             c("0"="Controls", "1"="Schizophrenia"))

d$Diagnosis # now it looks right

# Tell the dataset what the different non-numeric parameters are (ID and Diagnosis are factors and not numeric)
d <- d %>%
  mutate(
    ID = as.factor(ID),
    Diagnosis = as.factor(Diagnosis)
  )

# Define the formula - designing the model. We have a between-study design. 
AltercentricDiagnosis_f0 <- bf( # f stands for formula. bf stands for "Bayesian formula)
  AltercentricIntrusion ~ 1 + Diagnosis # when we say 1 we only use controls (baseline) as an intercept
)

AltercentricDiagnosis_f <- bf(
  AltercentricIntrusion ~ 0 + Diagnosis # when using 0 it gives us the estimate for both controls and schizophrenia
)

# Design the priors
priors_f0 <- get_prior(AltercentricDiagnosis_f0, family = gaussian, d) # The family = Guassian, because the outcome is a continuous variable
priors_f <- get_prior(AltercentricDiagnosis_f, family = gaussian, d)

# We see that we need a beta prior and a sigma prior

# We need two different priors: beta and sigma.
priorDiagnosis <- c(
  prior(normal(4, 1), class = b), # beta is the expectation of the difference between schizophrenia and controls. We say that the beta is normally distributed, and that the mean is 4 and the standard deviation is 1.
  
  prior(normal(1, 2), class = sigma) # sigma is the average error that we expect. 
) 

# Test the priors. We want to check whether the priors make any sense.
AltercentricDiagnosis_PriorCheck_m <- brm( # m stands for model - brm() makes a Bayesian regression model
  formula = AltercentricDiagnosis_f,
  data = d,
  family = gaussian, # Gaussian because the outcome is continuous
  prior = priorDiagnosis,
  sample_prior = "only", # we sample the prior in order to test the prior
  file = "AltercentricDiagnosis_PPC" # saving the model, so we do not have to run it every time we open R
)

# We check what the predictions look like given the prior and not the data. We set the number of simulations to 100.
pp_check(AltercentricDiagnosis_PriorCheck_m, nsamples = 100)

# What we see is that the prior has a very long tail. In order to fix this we make a prior for the sigma - in order to expect error

# Now that we have made a prior for sigma as well, we can fit the model

## Fitting the model
AltercentricDiagnosis_m <- brm(
  formula = AltercentricDiagnosis_f,
  data = d,
  family = gaussian,
  prior = priorDiagnosis,
  sample_prior = T,
  file = "AltercentricDiagnosis_model"
)

# Posterior predictive check. We want to look at whether the posterior has learned from the prior, which is what we expect when we look at the posterior predictive check
pp_check(AltercentricDiagnosis_m, nsamples = 100, replace = T)
# The data looks good!
# The light blue is the prior for the difference between schizophrenia and control (it is very spread, which means that it is very uncertain). 
# The dark blue is the posterior (this is much more certain - must less variance), which tells us that it has actually learned from the data, and makes more confident predictions. 

# Conclusion: the prior is not off, and the posterior has learned from the data. This is good!

## Check the model for warnings. 
AltercentricDiagnosis_m
# We see that the family is Gaussian, that we have two parameters (my and sigma), two estimates (one for control and one for schizophrenia), standard errors, diagnostic (1s and high numbers are good), the sigma is above 1 which makes perfect sense. 

# Hypothesis testing + updating check. We want to compare the prior and the posterior - checking if the model has learned anything. 
plot(hypothesis(AltercentricDiagnosis_m,
           "DiagnosisSchizophrenia > DiagnosisControls"))
# The function hypothesis() tests a specific hypothesis, which we specify. We check whether the diagnosis for schizoprehina is higher than the diagnosis for control 

# We see that the hypothesis that schizophrenia has a higher altercentric intrusion than controls is correct. 
# The evidence ratio is very high (the evidence ratio says that if we count how many of the samples in the posteiror are compatible with the hypothesis (that the altercentric intrusion is higher for schizophrenia than for controls) what is the ratio between the samples - we see that the ratio is that there are many more samples above 0 compared to at 0 and below). Thus, there is evidence given the data and the model, that shcizophreina has a higher level of altercentric intrusion compared to controls

hypothesis(AltercentricDiagnosis_m,
           "DiagnosisSchizophrenia > DiagnosisControls")

conditional_effects(AltercentricDiagnosis_m)

# We can plot the results: the prior for the hypothesis
plot(conditional_effects(AltercentricDiagnosis_m), points=T)

# What we see is a plot of the estimates for the schizoprehics and for controls. This is what the true data looks like. If we only look at one individual we do not have a high predictiability of whether they have schizoprhenia just by looking at altercentric intrusion, but at the population level there is a clear effect of altercentric intrusion on schizophrenia diagnosis.  

# Answer to the question for part 1: Yes, schizophrenia involves altercentric intrusion.

```

Reporting the model: 
The model indicates a credible differnce in altercentric intrusion in the two groups supporting our hypothesis (beta = 0.36, CIs = 0.16, 0.56, ER = 499). Controls show on average altercentric intrusion effect of 3.86 (CIs = 3.74, 3.98), and schizophrenia show altercentric intrusion effect of 4.22 (CIs = 4.01, 4.43). [Add plot of the effects]. The model had no divergences (as displayed by the warnings), A Rhat of 1, and Effective sample sizes of above 2000 for both Bulk and Tail. [Add prior and posterior checks plots; add updating check plot]

Q1.2) Is altercentric intrusion related to specific symptoms *in the patients*? Identify which of the symptoms could be relevant. Should you include more than one symptom? Build models, priors, predictive checks. Assess the evidence and report models and results, including plots. Discuss whether the results make sense.

```{r}
# First we make a subset of the data only containing the data for the patients (schizophrenics)
schizophrenia_data <- dplyr::filter(d, Diagnosis == "Schizophrenia")

# Now we scale all the variables
schizophrenia_data <- schizophrenia_data %>% 
  mutate(
    AltercentricIntrusion = scale(AltercentricIntrusion),
    VoiceHearing = scale(VoiceHearing),
    MindReading = scale(MindReading),
    Apathy = scale(Apathy)
  )

# Now we need to make a model predicting altercentric intrusion from each predictor (apathy, voice hearing, and mindreading). Thus we need to make three models.

```

```{r}
# VOICE HEARING #
summary(schizophrenia_data$VoiceHearing) # checking which mean and standard deviation we need for the prior

voicehearing_f <- bf( # making a bayesian formula predicting altercentric intrusion from voice hearing
  AltercentricIntrusion ~ 1 + VoiceHearing
)

# We check which priors we need 
priors_f_voicehearing <- get_prior(voicehearing_f, family = gaussian, schizophrenia_data)

# Making a prior for the beta and the sigma - the values come from the summary of voice hearing
prior_for_voicehearing <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b),
  prior(normal(1, 2), class = sigma) 
) 

voicehearing_PriorCheck_m <- brm( # making a bayesian regression model
  formula = voicehearing_f,
  data = schizophrenia_data,
  family = gaussian, 
  prior = prior_for_voicehearing,
  sample_prior = "only",
  file = "VoiceHearing_PPC_Final"
)

# We check our prior and whether they look sensible or not
pp_check_prior_voicehearing <- pp_check(voicehearing_PriorCheck_m, nsamples = 100) 
pp_check_prior_voicehearing
# The priors look fine

# Now that we have checked our priors we can make the model
voicehearing_model <- brm(
  formula = voicehearing_f,
  data = schizophrenia_data,
  family = gaussian,
  prior = prior_for_voicehearing,
  sample_prior = T,
  file = "VoiceHearing_model_final"
)

# We check whether the posterior has learned from the prior
pp_check_posterior_voicehearing <- pp_check(voicehearing_model, nsamples = 100)
pp_check_posterior_voicehearing

# Checking whether the model gives us any warning.
voicehearing_model

# Assessing the model
summary(voicehearing_model)

# Now we can test the hypothesis that effect of voicehearing on altercentric instrusion is above 0
plot(hypothesis(voicehearing_model, "VoiceHearing > 0"))

hypothesis(voicehearing_model,"VoiceHearing > 0")

conditional_effects(voicehearing_model)

plot(conditional_effects(voicehearing_model), points=T)

```

```{r}
# APATHY #
summary(schizophrenia_data$Apathy)

Apathy_f <- bf(
  AltercentricIntrusion ~ 1 + Apathy 
)

priors_f_Apathy <- get_prior(Apathy_f, family = gaussian, schizophrenia_data)

prior_for_Apathy <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b),
  prior(normal(1, 2), class = sigma) 
) 

Apathy_PriorCheck_m <- brm( 
  formula = Apathy_f,
  data = schizophrenia_data,
  family = gaussian, 
  prior = prior_for_Apathy,
  sample_prior = "only",
  file = "Apathy_PPC_final"
)

pp_check_prior_Apathy <- pp_check(Apathy_PriorCheck_m, nsamples = 100)
pp_check_prior_Apathy

Apathy_model <- brm(
  formula = Apathy_f,
  data = schizophrenia_data,
  family = gaussian,
  prior = prior_for_Apathy,
  sample_prior = T,
  file = "Apathy_model_final"
)

pp_check_posterior_Apathy <- pp_check(Apathy_model, nsamples = 100) 
pp_check_posterior_Apathy

Apathy_model

summary(Apathy_model)

# Now we can test the hypothesis that effect of apathy on altercentric instrusion is above 0
plot(hypothesis(Apathy_model, "Apathy > 0"))

hypothesis(Apathy_model,"Apathy > 0")

conditional_effects(Apathy_model)

plot(conditional_effects(Apathy_model), points=T)

```

```{r}
# MIND READING #
summary(schizophrenia_data$MindReading)

MindReading_f <- bf(
 AltercentricIntrusion ~ 1 + MindReading 
)

priors_f_MindReading <- get_prior(MindReading_f, family = gaussian, schizophrenia_data)

prior_for_MindReading <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b),
  prior(normal(1, 2), class = sigma) 
) 

MindReading_PriorCheck_m <- brm( 
  formula = MindReading_f,
  data = schizophrenia_data,
  family = gaussian, 
  prior = prior_for_MindReading,
  sample_prior = "only",
  file = "MindReading_PPC_final1"
)

pp_check_prior_MindReading <- pp_check(MindReading_PriorCheck_m, nsamples = 100) 
pp_check_prior_MindReading

MindReading_model <- brm(
  formula = MindReading_f,
  data = schizophrenia_data,
  family = gaussian,
  prior = prior_for_MindReading,
  sample_prior = T,
  file = "MindReading_model_final1"
)

pp_check_posterior_MindReading <- pp_check(MindReading_model, nsamples = 100) 
pp_check_posterior_MindReading

MindReading_model

summary(MindReading_model)

# Now we can test the hypothesis that effect of mind reading on altercentric instrusion is above 0
plot(hypothesis(MindReading_model, "MindReading > 0"))

hypothesis(MindReading_model,"MindReading > 0")

conditional_effects(MindReading_model)

plot(conditional_effects(MindReading_model), points=T)

```

## MULTIVARIATE MODELS ##
```{r}
# Now that we have made a model for each predictor, we want to make models that include more than one predictor ata time. 

# APATHY AND VOICE HEARING #
summary(schizophrenia_data$Apathy)
summary(schizophrenia_data$VoiceHearing)

multivariateModel_f <- bf(AltercentricIntrusion ~ 1 + Apathy + VoiceHearing)

priors_multivariate <- get_prior(multivariateModel_f, family = gaussian, schizophrenia_data)

prior_for_multivariateModel <- c(
  prior(normal(0,1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = Apathy),
  prior(normal(0, 0.3), class = b, coef = VoiceHearing),
  prior(normal(1, 2), class = sigma)
)

multivariateModel_PriorCheck_m <- brm( 
  formula = multivariateModel_f,
  data = schizophrenia_data,
  family = gaussian, 
  prior = prior_for_multivariateModel,
  sample_prior = "only",
  file = "ApathyVoiceHearing_PPC_final1"
)

pp_check_prior_multivariateModel <- pp_check(multivariateModel_PriorCheck_m, nsamples = 100) 
pp_check_prior_multivariateModel

multivariate_model <- brm(
  formula = multivariateModel_f,
  data = schizophrenia_data,
  family = gaussian,
  prior = prior_for_multivariateModel,
  sample_prior = T,
  file = "ApathyVoiceHearing_model_final1"
)

pp_check_posterior_multivariate <- pp_check(multivariate_model, nsamples = 100) 
pp_check_posterior_multivariate

multivariate_model

summary(multivariate_model)

plot(hypothesis(multivariate_model, "Apathy > 0"))
plot(hypothesis(multivariate_model, "VoiceHearing > 0"))

hypothesis(multivariate_model,"Apathy > 0")
hypothesis(multivariate_model,"VoiceHearing > 0")

conditional_effects(multivariate_model)

plot(conditional_effects(multivariate_model), points=T)

```

```{r}
# APATHY AND MIND READING 
summary(schizophrenia_data$Apathy)
summary(schizophrenia_data$MindReading)

multivariateModel2_f <- bf(AltercentricIntrusion ~ 1 + Apathy + MindReading)

priors_multivariate2 <- get_prior(multivariateModel2_f, family = gaussian, schizophrenia_data)

prior_for_multivariateModel2 <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = Apathy),
  prior(normal(0, 0.3), class = b, coef = MindReading),
  prior(normal(1, 2), class = sigma)
)

multivariateModel2_PriorCheck_m <- brm( 
  formula = multivariateModel2_f,
  data = schizophrenia_data,
  family = gaussian, 
  prior = prior_for_multivariateModel2,
  sample_prior = "only",
  file = "ApathyMindReading_PPC_final"
)

pp_check_prior_multivariateModel2 <- pp_check(multivariateModel2_PriorCheck_m, nsamples = 100) 
pp_check_prior_multivariateModel2

multivariate_model2 <- brm(
  formula = multivariateModel2_f,
  data = schizophrenia_data,
  family = gaussian,
  prior = prior_for_multivariateModel2,
  sample_prior = T,
  file = "ApathyMindReading_model_final"
)

pp_check_posterior_multivariate2 <- pp_check(multivariate_model2, nsamples = 100) 
pp_check_posterior_multivariate2

multivariate_model2

summary(multivariate_model2)

plot(hypothesis(multivariate_model2, "Apathy > 0"))
plot(hypothesis(multivariate_model2, "MindReading > 0"))

hypothesis(multivariate_model2,"Apathy > 0")
hypothesis(multivariate_model2,"MindReading > 0")

conditional_effects(multivariate_model2)

plot(conditional_effects(multivariate_model2), points=T)

```

```{r}
# VOICE HEARING AND MIND READING # 
multi_3_f <- bf(
  AltercentricIntrusion ~ 1 + MindReading + VoiceHearing
)

priors_f_multi_3 <- get_prior(multi_3_f, family = gaussian, schizophrenia_data)

prior_for_multi_3 <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = MindReading), 
  prior(normal(0, 0.3), class = b, coef = VoiceHearing),
  prior(normal(1, 2), class = sigma) 
) 

multi_3_PriorCheck_m <- brm( 
  formula = multi_3_f,
  data = schizophrenia_data,
  family = gaussian, 
  prior = prior_for_multi_3,
  sample_prior = "only",
  file = "multi_3_PriorCheck_m_final"
)

pp_check_prior_multi_3 <- pp_check(multi_3_PriorCheck_m, nsamples = 100) # checking whether the prior looks sensible

multi_m_3 <- brm(
  formula = multi_3_f,
  data = schizophrenia_data,
  family = gaussian,
  prior = prior_for_multi_3,
  sample_prior = T,
  file = "multi_m_3_final"
)

pp_check_posterior_multi_3 <- pp_check(multi_m, nsamples = 100) # checking whether the posterior has learned from the prior
pp_check_prior_multi_3
pp_check_posterior_multi_3

multi_m_3

summary(multi_m_3)

plot(hypothesis(multi_m_3,"MindReading > 0"))
plot(hypothesis(multi_m_3,"VoiceHearing > 0"))

hypothesis(multi_m_3,"MindReading > 0")
hypothesis(multi_m_3,"VoiceHearing > 0")

conditional_effects(multi_m_3)
plot(conditional_effects(multi_m_3), points=T)

```

```{r}
# MIND READING, VOICE HEARING, APATHY # 
all_f <- bf(
  AltercentricIntrusion ~ 1 + MindReading + VoiceHearing + Apathy
)

priors_f_all <- get_prior(all_f, family = gaussian, schizophrenia_data)

prior_for_all <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = Apathy),
  prior(normal(0, 0.3), class = b, coef = MindReading), 
  prior(normal(0, 0.3), class = b, coef = VoiceHearing),
  prior(normal(1, 2), class = sigma) 
) 

all_PriorCheck_m <- brm( 
  formula = all_f,
  data = schizophrenia_data,
  family = gaussian, 
  prior = prior_for_all,
  sample_prior = "only",
  file = "all_PriorCheck_m_final"
)

pp_check_prior_all <- pp_check(all_PriorCheck_m, nsamples = 100) # checking whether the prior looks sensible

all_m <- brm(
  formula = all_f,
  data = schizophrenia_data,
  family = gaussian,
  prior = prior_for_all,
  sample_prior = T,
  file = "all_m_final"
)

pp_check_posterior_all <- pp_check(all_m, nsamples = 100) # checking whether the posterior has learned from the prior
pp_check_prior_all
pp_check_posterior_all

all_m

summary(all_m)

#plot(hypothesis(multi_m_3,"MindReading > 0"))
#plot(hypothesis(multi_m_3,"VoiceHearing > 0"))

hypothesis(all_m,"MindReading > 0")
hypothesis(all_m,"VoiceHearing > 0")
hypothesis(all_m,"Apathy > 0")
#conditional_effects(all_m)

# We can plot the results: the prior for the hypothesis
plot(conditional_effects(all_m), points=T)

```

# MODEL COMPARISON # 
```{r}
# Model Comparison: now we compare all of the models
AltercentricVoiceHearing_m <- add_criterion(voicehearing_model, criterion = "loo")
AltercentricMindReading_m <- add_criterion(MindReading_model, criterion = "loo")
AltercentricApathy_m <- add_criterion(Apathy_model, criterion = "loo")

Altercentric_Apathy_VoiceHearing <- add_criterion(multivariate_model, criterion = "loo")
Altercentric_Apathy_MindReading <- add_criterion(multivariate_model2, criterion = "loo")
Altercentric_MindReading_VoiceHearing <- add_criterion(multi_m_3, criterion = "loo")
Altercentric_MRVHA <- add_criterion(all_m, criterion = "loo")

loo_compare(AltercentricVoiceHearing_m,
            AltercentricMindReading_m,
            AltercentricApathy_m,
            Altercentric_Apathy_VoiceHearing,
            Altercentric_Apathy_MindReading,
            Altercentric_MindReading_VoiceHearing,
            Altercentric_MRVHA)
            
loo_model_weights(AltercentricVoiceHearing_m,
            AltercentricMindReading_m,
            AltercentricApathy_m,
            Altercentric_Apathy_VoiceHearing,
            Altercentric_Apathy_MindReading,
            Altercentric_MindReading_VoiceHearing,
            Altercentric_MRVHA)

```

## Second part

Q2.1) However, we know that the diagnosis is based on symptom assessment: if the overall sum of symptoms is severe enough, the participant gets a diagnosis. In other words, by selecting the patients, and including the symptoms in the model we might have inadvertently introduced an issue in our inference. Do try to draw a causal graph (Directed Acyclical Graph) of the variables and compare it with the types of causal graphs presented in the slides. Discuss which biases you might have introduced.

[answered in the document]

Q2.2.) Redesign your analysis following the graph and report how the results change

```{r}
# According to the DAG graph we have made we are introducing a bias by only including schizophrenic patients. Thus, we need to rerun all of the models with all participants (both healthy and schizophrenic)

# First we scale all variables
d <- d %>% 
  mutate(
    AltercentricIntrusion = scale(AltercentricIntrusion),
    VoiceHearing = scale(VoiceHearing),
    MindReading = scale(MindReading),
    Apathy = scale(Apathy)
  )

summary(d$VoiceHearing)
summary(d$Apathy)
summary(d$MindReading)
```

Now we make all of the models again with data that includes all participants:
```{r}
# VOICE HEARING #
voicehearing_f <- bf(AltercentricIntrusion ~ 1 + VoiceHearing)

priors_f_voicehearing <- get_prior(voicehearing_f, family = gaussian, data = d)

prior_for_voicehearing <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b),
  prior(normal(1, 2), class = sigma) 
) 

voicehearing_PriorCheck_m <- brm(
  formula = voicehearing_f,
  data = d,
  family = gaussian, 
  prior = prior_for_voicehearing,
  sample_prior = "only",
  file = "VoiceHearing_PPC_alldata"
)

pp_check_prior_voicehearing <- pp_check(voicehearing_PriorCheck_m, nsamples = 100) 
pp_check_prior_voicehearing

voicehearing_model <- brm(
  formula = voicehearing_f,
  data = d,
  family = gaussian,
  prior = prior_for_voicehearing,
  sample_prior = T,
  file = "VoiceHearing_model_alldata"
)

pp_check_posterior_voicehearing <- pp_check(voicehearing_model, nsamples = 100)
pp_check_posterior_voicehearing

voicehearing_model

summary(voicehearing_model)

plot(hypothesis(voicehearing_model, "VoiceHearing > 0"))

hypothesis(voicehearing_model,"VoiceHearing > 0")

conditional_effects(voicehearing_model)

plot(conditional_effects(voicehearing_model), points=T)

```

```{r}
# APATHY #
Apathy_f <- bf(AltercentricIntrusion ~ 1 + Apathy )

priors_f_Apathy <- get_prior(Apathy_f, family = gaussian, d)

prior_for_Apathy <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b),
  prior(normal(1, 2), class = sigma) 
) 

Apathy_PriorCheck_m <- brm( 
  formula = Apathy_f,
  data = d,
  family = gaussian, 
  prior = prior_for_Apathy,
  sample_prior = "only",
  file = "Apathy_PPC_alldata"
)

pp_check_prior_Apathy <- pp_check(Apathy_PriorCheck_m, nsamples = 100)
pp_check_prior_Apathy

Apathy_model <- brm(
  formula = Apathy_f,
  data = d,
  family = gaussian,
  prior = prior_for_Apathy,
  sample_prior = T,
  file = "Apathy_model_alldata"
)

pp_check_posterior_Apathy <- pp_check(Apathy_model, nsamples = 100) 
pp_check_posterior_Apathy

Apathy_model

summary(Apathy_model)

plot(hypothesis(Apathy_model, "Apathy > 0"))

hypothesis(Apathy_model,"Apathy > 0")

conditional_effects(Apathy_model)

plot(conditional_effects(Apathy_model), points=T)

```

```{r}
# MIND READING #
MindReading_f <- bf(AltercentricIntrusion ~ 1 + MindReading)

priors_f_MindReading <- get_prior(MindReading_f, family = gaussian, d)

prior_for_MindReading <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b),
  prior(normal(1, 2), class = sigma) 
) 

MindReading_PriorCheck_m <- brm( 
  formula = MindReading_f,
  data = d,
  family = gaussian, 
  prior = prior_for_MindReading,
  sample_prior = "only",
  file = "MindReading_PPC_alldata"
)

pp_check_prior_MindReading <- pp_check(MindReading_PriorCheck_m, nsamples = 100) 
pp_check_prior_MindReading

MindReading_model <- brm(
  formula = MindReading_f,
  data = d,
  family = gaussian,
  prior = prior_for_MindReading,
  sample_prior = T,
  file = "MindReading_model_alldata"
)

pp_check_posterior_MindReading <- pp_check(MindReading_model, nsamples = 100) 
pp_check_posterior_MindReading

MindReading_model

summary(MindReading_model)

plot(hypothesis(MindReading_model, "MindReading > 0"))

hypothesis(MindReading_model,"MindReading > 0")

conditional_effects(MindReading_model)

plot(conditional_effects(MindReading_model), points=T)

```

# MULTIVARIATE MODELS # 
```{r}
# Now we can run the multivariate models with the data that includes btoh control and patients.

# APATHY AND VOICE HEARING ##
multivariateModel_f <- bf(AltercentricIntrusion ~ 1 + Apathy + VoiceHearing)

priors_multivariate <- get_prior(multivariateModel_f, family = gaussian, d)

prior_for_multivariateModel <- c(
  prior(normal(0,1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = Apathy),
  prior(normal(0, 0.3), class = b, coef = VoiceHearing),
  prior(normal(1, 2), class = sigma)
)

multivariateModel_PriorCheck_m <- brm( 
  formula = multivariateModel_f,
  data = d,
  family = gaussian, 
  prior = prior_for_multivariateModel,
  sample_prior = "only",
  file = "ApathyVoiceHearing_PPC_alldata"
)

pp_check_prior_multivariateModel <- pp_check(multivariateModel_PriorCheck_m, nsamples = 100) 
pp_check_prior_multivariateModel

multivariate_model <- brm(
  formula = multivariateModel_f,
  data = d,
  family = gaussian,
  prior = prior_for_multivariateModel,
  sample_prior = T,
  file = "ApathyVoiceHearing_model_alldata"
)

pp_check_posterior_multivariate <- pp_check(multivariate_model, nsamples = 100) 
pp_check_posterior_multivariate

multivariate_model

summary(multivariate_model)

plot(hypothesis(multivariate_model, "Apathy > 0"))
plot(hypothesis(multivariate_model, "VoiceHearing > 0"))

hypothesis(multivariate_model,"Apathy > 0")
hypothesis(multivariate_model,"VoiceHearing > 0")

conditional_effects(multivariate_model)

plot(conditional_effects(multivariate_model), points=T)

```

```{r}
# APATHY AND MIND READING #
multivariateModel2_f <- bf(AltercentricIntrusion ~ 1 + Apathy + MindReading)

priors_multivariate2 <- get_prior(multivariateModel2_f, family = gaussian, d)

prior_for_multivariateModel2 <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = Apathy),
  prior(normal(0, 0.3), class = b, coef = MindReading),
  prior(normal(1, 2), class = sigma)
)

multivariateModel2_PriorCheck_m <- brm( 
  formula = multivariateModel2_f,
  data = d,
  family = gaussian, 
  prior = prior_for_multivariateModel2,
  sample_prior = "only",
  file = "ApathyMindReading_PPC_alldata"
)

pp_check_prior_multivariateModel2 <- pp_check(multivariateModel2_PriorCheck_m, nsamples = 100) 
pp_check_prior_multivariateModel2

multivariate_model2 <- brm(
  formula = multivariateModel2_f,
  data = d,
  family = gaussian,
  prior = prior_for_multivariateModel2,
  sample_prior = T,
  file = "ApathyMindReading_model_alldata"
)

pp_check_posterior_multivariate2 <- pp_check(multivariate_model2, nsamples = 100) 
pp_check_posterior_multivariate2

multivariate_model2

summary(multivariate_model2)

plot(hypothesis(multivariate_model2, "Apathy > 0"))
plot(hypothesis(multivariate_model2, "MindReading > 0"))

hypothesis(multivariate_model2,"Apathy > 0")
hypothesis(multivariate_model2,"MindReading > 0")

conditional_effects(multivariate_model2)

plot(conditional_effects(multivariate_model2), points=T)

```

```{r}
# VOICE HEARING AND MIND READING # 
multi_3_f <- bf(
  AltercentricIntrusion ~ 1 + MindReading + VoiceHearing
)

priors_f_multi_3 <- get_prior(multi_3_f, family = gaussian, d)

prior_for_multi_3 <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = MindReading), 
  prior(normal(0, 0.3), class = b, coef = VoiceHearing),
  prior(normal(1, 2), class = sigma) 
) 

multi_3_PriorCheck_m <- brm( 
  formula = multi_3_f,
  data = d,
  family = gaussian, 
  prior = prior_for_multi_3,
  sample_prior = "only",
  file = "multi_3_PriorCheck_m_alldata"
)

pp_check_prior_multi_3 <- pp_check(multi_3_PriorCheck_m, nsamples = 100) # checking whether the prior looks sensible

multi_m_3 <- brm(
  formula = multi_3_f,
  data = d,
  family = gaussian,
  prior = prior_for_multi_3,
  sample_prior = T,
  file = "multi_m_3_final_all_data"
)

pp_check_posterior_multi_3 <- pp_check(multi_m, nsamples = 100) # checking whether the posterior has learned from the prior
pp_check_posterior_multi_3

multi_m_3

summary(multi_m_3)

plot(hypothesis(multi_m_3,"MindReading > 0"))
plot(hypothesis(multi_m_3,"VoiceHearing > 0"))

hypothesis(multi_m_3,"MindReading > 0")
hypothesis(multi_m_3,"VoiceHearing > 0")
conditional_effects(multi_m_3)

plot(conditional_effects(multi_m_3), points=T)

```

```{r}
# APATHY, VOICE HEARING, MIND READING # 
all_f <- bf(
  AltercentricIntrusion ~ 1 + MindReading + VoiceHearing + Apathy
)

priors_f_all <- get_prior(all_f, family = gaussian, d)

prior_for_all <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b, coef = Apathy),
  prior(normal(0, 0.3), class = b, coef = MindReading), 
  prior(normal(0, 0.3), class = b, coef = VoiceHearing),
  prior(normal(1, 2), class = sigma) 
) 

all_PriorCheck_m <- brm( 
  formula = all_f,
  data = d,
  family = gaussian, 
  prior = prior_for_all,
  sample_prior = "only",
  file = "all_PriorCheck_m_alldata"
)

pp_check_prior_all <- pp_check(all_PriorCheck_m, nsamples = 100)

all_m <- brm(
  formula = all_f,
  data = d,
  family = gaussian,
  prior = prior_for_all,
  sample_prior = T,
  file = "all_m_final_alldata"
)

pp_check_posterior_all <- pp_check(all_m, nsamples = 100) # checking whether the posterior has learned from the prior
pp_check_prior_all
pp_check_posterior_all

all_m

summary(all_m)

plot(hypothesis(multi_m_3,"MindReading > 0"))
plot(hypothesis(multi_m_3,"VoiceHearing > 0"))
hypothesis(all_m,"MindReading > 0")
hypothesis(all_m,"VoiceHearing > 0")
hypothesis(all_m,"Apathy > 0")
plot(conditional_effects(all_m), points=T)

```

# MODEL COMPARISON # 
```{r}
# Now we compare all of the models that use all of the data
AltercentricVoiceHearing_m <- add_criterion(voicehearing_model, criterion = "loo")
AltercentricMindReading_m <- add_criterion(MindReading_model, criterion = "loo")
AltercentricApathy_m <- add_criterion(Apathy_model, criterion = "loo")

Altercentric_Apathy_VoiceHearing <- add_criterion(multivariate_model, criterion = "loo")
Altercentric_Apathy_MindReading <- add_criterion(multivariate_model2, criterion = "loo")
Altercentric_MindReading_VoiceHearing <- add_criterion(multi_m_3, criterion = "loo")
Altercentric_MRVHA <- add_criterion(all_m, criterion = "loo")

loo_compare(AltercentricVoiceHearing_m,
            AltercentricMindReading_m,
            AltercentricApathy_m,
            Altercentric_Apathy_VoiceHearing,
            Altercentric_Apathy_MindReading,
            Altercentric_MindReading_VoiceHearing,
            Altercentric_MRVHA)
            
loo_model_weights(AltercentricVoiceHearing_m,
            AltercentricMindReading_m,
            AltercentricApathy_m,
            Altercentric_Apathy_VoiceHearing,
            Altercentric_Apathy_MindReading,
            Altercentric_MindReading_VoiceHearing,
            Altercentric_MRVHA)

```

## Third part

These issues are very difficult to think through, and not knowing the causal mechanisms generating the data in advance makes our inferences even more unreliable. To explore these issues, I recommend using simulations. In other words, defining a "true" model, generating data from it and assessing what different analyses would lead you to infer (and therefore which biases they might introduce). You can find the code I used to simulate your data below.

Q3.1) Look through the code and identify whether the results you have match the underlying truth. Discuss what you have learned.



Q3.2) OPTIONAL: is this a general pattern? Try varying the parameters (e.g. correlation values) and assess whether the new dataset(s) leads to the same biases in your analysis.
```{r}
pacman::p_load(MASS, tidyverse, psych)

seed <- 1981 # Defining a seed so the results are always the same
n <- 300 # Defining the amount of participants

SymptomCorr <- .2 # Defining the correlation of symptoms (as they tend to co-occur)
EffectCorrRel <- .2 # Defining the correlation between relevant symptoms and effect (Some symptoms are positively correlated with the effect)
EffectCorrIrrel <- 0 # Defining the correlation between irrelevant symptoms and effect (none)

# Creating the variance-covariance matrix for the variables we want to generate (3 symptoms, 1 effect)
Sigma <- matrix(data=c(1,SymptomCorr,SymptomCorr,EffectCorrRel,
                       SymptomCorr,1,SymptomCorr,EffectCorrRel,
                       SymptomCorr,SymptomCorr,1,EffectCorrIrrel,
                       EffectCorrRel,EffectCorrRel,EffectCorrIrrel,1),
                       nrow=4,ncol=4)

## Generate data from a multivariate (mvr) normal (n) distribution
d <- mvrnorm(n = n, # number of participant
        mu = c(1.2, 1.2, 1.2, 4), # mean of each variable
        Sigma) # variance co-variance matrix

# Giving meaningful names to variables and add ID
d <- data.frame(
  VoiceHearing = d[,1], 
  MindReading =  d[,2],
  Apathy =  d[,3], 
  AltercentricIntrusion = d[,4],
  ID = seq(nrow(d)))

# Assessing whether the participant has schizophrenia (high enough sum of symptoms)
# Here we choose participants scoring above 75% percentile (the most severe ones)
d$Diagnosis <- 0
d$Diagnosis[(d$VoiceHearing + d$MindReading + d$Apathy) > 
              quantile(d$VoiceHearing + d$MindReading + d$Apathy, .75)] <-1

## Plotting the relation between variables in schizophrenia
d1 <- d %>% subset(Diagnosis==1) %>% dplyr::select(-Diagnosis, -ID)
pairs.panels(d1)

## Plotting the relation between variables all participants
pairs.panels(dplyr::select(d,-Diagnosis, -ID))

write_csv(d, "data/Ass3.csv")
```

